---
title: "EXPERIMENT 4_bikeshare data "
output: html_notebook
---

# Data Packages 

Baseine pacakges for our analysis

```{r}
library(tidyverse)
library(janitor)
library(here)
library(openmeteo)
```


# Loading the Bikeshare Raw Data
```{r}
df1 <- read_csv("data_raw/202309-capitalbikeshare-tripdata.csv", show_col_types = FALSE) %>% clean_names()
```

# Creating Dataframe 2 - Interjoing 
```{r}
df2s=df1 %>% 
  select(rideable_type,member_casual,
                	contains("start"),ride_id)%>% 
  mutate(start_stop="start")%>%
  rename(t=started_at,
     	station_name=start_station_name,
     	station_id=start_station_id,
     	lat=start_lat,
     	lng=start_lng)

df2e=df1 %>%
select(ride_id,rideable_type,member_casual,
                	contains("end")) %>%
  mutate(start_stop="stop") %>%
  rename(t=ended_at,
     	station_name=end_station_name,
     	station_id=end_station_id,
     	lat=end_lat,
     	lng=end_lng)

df2=bind_rows(df2s,df2e) %>%
  arrange(t) %>%
  mutate(rider_delta=(start_stop=="start")*2-1) %>% #change in ridership 
  mutate(riders=cumsum(rider_delta)) %>%
  relocate(riders,.after=t)

```

# Plotting the Dataframe

```{r}
df2 %>% 
  ggplot(aes(t,riders)) +
  geom_line()
```
# Creating a subsampled dataset 

```{r}
df_s=df2 %>% slice_head(n=1000)

```

Why looking at slicing to every one hundredth data point will be bad? - Will likely track features however will jump data points whether the data is sparse or not. Want the data to be equally spaced in time.

Starting with 100 rows to get the calculation right and then see if oyu can do it with a bigger data set. 
Data stored in number of seconds based on a reference point


Round down to the nearest 10 mins, relocate t_f next to t, and filter the data points by t_f

```{r}
df_e=df_s |>
  mutate(t_f=floor_date(t, "10 mins")) %>%
   relocate(t_f,.after=t) %>%
  slice_head(n=1,by=t_f)
```


# Applying to Previous Manipulations to the Entire Data Set 
```{r}
df_r=df2 |>
  mutate(t_f=floor_date(t,"10 mins")) %>%
  relocate(t_f,.after=t) %>%
  slice_head(n=1,by=t_f)
```

 
# Creating the Associated Plot 
```{r}

p1=df2 %>% 
  filter(day(t)==18) %>%
  ggplot(aes(x=t, y=riders)) +
  geom_line() +
  ggtitle("Riders on 18Sep")

p1+
  geom_line(data=df_r %>% filter(day(t)==18),
  color="red")

```
 

```{r}
df_r=df2 |>
  mutate(t_f=floor_date(t,"1 mins")) %>%
  relocate(t_f,.after=t) %>%
  slice_head(n=1,by=t_f)

```


```{r}
p1=df2 %>% 
  filter(day(t)==18) %>%
  ggplot(aes(x=t, y=riders)) +
  geom_line() +
  ggtitle("Riders on 18Sep")

p1+
  geom_line(data=df_r %>% filter(day(t)==18),
  color="red")
```

% Get Weather Data
```{r}
df_w=weather_history("Washington",
                	start = "2023-09-01",
                	end = "2023-09-30",
                	hourly = c("apparent_temperature",
                           	"wind_speed_10m",
                           	"precipitation")
)

```
# Merging bike and weather data

```{r}
df_s=df2 %>% slice_sample(n=1000)
df_j=df_s %>% left_join(df_w,
                     	by=join_by(closest(t>=datetime)))
```


```{r}
df_j=df_s %>% 
   left_join(df_w,by=join_by(closest(t>=datetime)))  %>%
   relocate(datetime, .after=t)
 
head(df_j)
```

## Investigating the Time Zone Mismatch Between the Bikeshare and Weather Datasets 

```{r}
df_j$t[1:5]
df_j$datetime[1:5]

```

```{r}
Sys.timezone()
```

```{r}
df_s_est <- force_tz(df_s, tzone = "America/New_York")
```


```{r}
df_j_est=df_s_est %>% 
   left_join(df_w,by=join_by(closest(t>=datetime)))  %>%
   relocate(datetime, .after=t)
 
head(df_j_est)
```

```{r}
df_j_est$t[1:5]
df_j_est$datetime[1:5]
```
```{r}
df2$t[1:5]
force_tz(df2$t[1:5],"America/New_York")
```

```{r}
df2c=df2 %>% mutate(t=force_tz(t,tzone="America/New_York")) #corrected
 
df_s2=df2c %>% slice_sample(n=1000)
 
df_j2=df_s2 %>% 
   left_join(df_w,by=join_by(closest(t>=datetime)))  %>%
   relocate(datetime, .after=t)
 
head(df_j2)

```

```{r}
dfc=df2c %>% 
  left_join(df_w,by=join_by(closest(t>=datetime))) %>%
   relocate(datetime, .after=t) %>%
  rename(atemp=hourly_apparent_temperature,
         wind=hourly_wind_speed_10m,
         prec=hourly_precipitation)
```


## A Visualization of the Data

Coloring when the rain is greater than 1- color each point based on whether it was raining or not 

```{r}
p2=dfc %>%
  ggplot(aes(x=t,y=riders,color=prec>1)) +
  geom_point()
p2
  
```


```{r}
p3 = dfc %>%
  filter(day(t)==23) %>%
  ggplot(aes(t,riders,color=wind))+
  geom_point() +
  ggtitle("Riders vs. Precipitation on Sep 10")

p3

```


```{r}
plotly::ggplotly(p3)
```







